---
Title: Intro Plumber
Date: 2018-10-12
Category: R
Tags: r, api, machine learning
slug: intro_plumber
Authors: Juan Camilo Orduz
Summary: In this post we present a simple example of how to expose a prediction model to a web API using the Plumber package. 
---

In this post we explore the basics of the [Plumber](https://www.rplumber.io) package. Our aim is to ilustrate how to fit a \\(L^2\\)-regularized linear model and expose it to a web API so that we can request predictions. 

## Prepare Notebook

Let us load the necessary libraries. 

```{r}
library(caret)
library(httr)
library(magrittr)
library(plumber)
library(tidyverse)
```

## Load Data

As a toy example we consider the `mtcars` data set. 

```{r}
df <- mtcars %>% as_tibble()

df %>% head
```

We want to fit a simple linear model to predict the variable `mpg`. 

*Warning:* We are not interested to find the "best model". Better feature engineering and hyperparameter tunig is not developed here because it is not the main purpose. 

## Correlation Plot 

For variable selection we consider a correlation plot. 

```{r}
df %>% cor %>% corrplot::corrplot()
```

From the visualization we see that the variables `wt`, `qsec` and `am` could be good predictors.

## Define and Train Model

We are going to use the [Caret](https://topepo.github.io/caret/index.html) package.

### Split Data 

```{r}
set.seed(seed = 0)

# Define observation matrix. 
X <- df %>% select(wt, qsec, am)
# Define target vector.
y <- df %>% pull(mpg)

# Define a partition of the data. 
partition <- createDataPartition(y = y, p = 0.75, list = FALSE) 

# Split the data into a training and test set. 
df.train <- df[partition, ]
df.test <- df[- partition, ]

X.train <- df.train %>% select(wt, qsec, am)
y.train <- df.train %>% pull(mpg)

X.test <- df.test %>% select(wt, qsec, am)
y.test <- df.test %>% pull(mpg)
```

### Data Pre-Processing

As we want to use a linear model, we neet to scale the variables. 

```{r}
# Define scaler object. 
scaler.obj <- preProcess(x = X.train, method = c('center', 'scale'))

# Transform the data. 
X.train.scaled <- predict(object = scaler.obj, newdata = X.train)
X.test.scaled <- predict(object = scaler.obj, newdata = X.test)
```

### Train Model

We fit \\(L^2\\)-regularization linear model using a 3-fold cross-validation to tune the regularization hyperparameter. 

```{r}
model.obj <-  train(x = X.train.scaled,
                    y = y.train,
                    method = 'ridge',
                    trControl = trainControl(method = 'cv', number = 3), 
                    metric = 'RMSE')
```

### Model Evaluation

Let us evaluate the model perforance. 

On the training set:

```{r}
model.obj$results %>% select(RMSE)
```

On the test set:

```{r}
y.pred <- predict(model.obj, newdata = X.test.scaled)

RMSE(pred = y.pred, obs = y.test)
```

The model seems to be stable. 

### Visualization

```{r}
tibble(y_test = y.test, y_pred = y.pred) %>% 
  ggplot() + 
  theme_light() + 
  geom_point(mapping = aes(x = y_test, y = y_pred)) + 
  geom_smooth(mapping = aes(x = y_test, y = y_pred, colour = 'y_pred ~ y_test'), method = 'lm', formula = y ~ x) + 
  geom_abline(mapping = aes(slope = 1, intercept = 0, colour = 'y_pred = y_test'), show.legend = TRUE) +
  ggtitle(label = 'Model Evaluation')
```

## Save Model

### Data Pipeline

We define a function which transforms and predicts for new incoming data. 

```{r}
GetNewPredictions <- function(model, transformer, newdata){
  
  newdata.transformed <- predict(object = transformer, newdata = newdata)
  
  new.predictions <- predict(object = model, newdata = newdata.transformed)
  
  return(new.predictions)
  
}
```

### Save Output Object 

```{r}
# Define Output object.
model.list <- vector(mode = 'list')
# Save scaler object.
model.list$scaler.obj <- scaler.obj
# Save fitted model.
model.list$model.obj <- model.obj
# Save transformation function. 
model.list$GetNewPredictions <- GetNewPredictions

saveRDS(object = model.list, file = 'model_list.rds')
```

## Write Plumber Script

This is the basic structure of a Plumber script. 

```{r, eval=FALSE}
# plumber.R

# Read model data.
model.list <- readRDS(file = 'model_list.rds')

#* @param wt
#* @param qsec
#* @param am
#* @get /predict
CalculatePrediction <- function(wt, qsec, am){
  
  wt %<>% as.numeric
  qsec %<>% as.numeric
  am %<>% as.numeric
  
  X.new <- tibble(wt = wt, qsec = qsec, am = am)
  
  y.pred <- GetNewPredictions(model = model.obj, transformer = scaler.obj, newdata = X.new)
  
  return(y.pred)
  
}
```

## Expose to API

To expose the model and get predictions we run:

```{r, eval=FALSE}
setwd(dir = here::here())

r <- plumb(file = 'plumber.R')

r$run(port = 8000)
```

We can use a `GET` request to obtain predictions. 

```{r, eval=FALSE}
GET(url = 'http://localhost:8000/predict?am=1&qsec=16.46&wt=2.62') %>% content
```

[[1]]
[1] 22.4974
